\mytitle{}

### 2019-07-31 paper/infocom.tex
Just the conclusion + new results to be handled...

\hypersetup{
	colorlinks,
	citecolor=black,
	filecolor=black,
	linkcolor=black,
	urlcolor=black,
	pdftitle={\mytitle{}},
%	pdfauthor={Kyle A. Simpson, Simon Rogers, Dimitrios P. Pezaros}
	pdfauthor={Anonymous Giraffe, Anonymous Badger, Anonymous Ibex}
}
\newcommand*{\email}[1]{\href{mailto:#1}{\nolinkurl{#1}} }

@@

\title{\mytitle{}}
\author{
%	Kyle~A.~Simpson\thanks{This work was supported by the Engineering and Physical Sciences Research Council [grant number EP/M508056/1].},
%	Kyle~A.~Simpson,
%	Simon~Rogers,
%	Dimitrios~P.~Pezaros,~\IEEEmembership{Senior~Member,~IEEE,},\\
%	School of Computing Science, University of Glasgow, Glasgow, G12 8RZ, UK\\
%	University of Glasgow, Glasgow, Scotland,\\
%	\email{k.simpson.1@research.gla.ac.uk}
Anonymous Giraffe,
Anonymous Badger,
Anonymous Ibex\\
Unnamed Department, Nowhere\\
\email{giraffe.a@unnamed.com}
}

@@

Reinforcement learning (RL) agents manage and monitor system health in an online manner.
Unsupervised, online learning is a key part of overcoming this detection problem.
We present the architecture of a system designed to host such RL agents in a typical network.
Furthermore, we introduce two agents which advance the state-of-the-art in using RL for DDoS mitigation.
Both of these act per flow, in a protocol-agnostic manner for any network topology.
This is supported by an in-depth investigation of feature suitability and empirical evaluation.

Our results show the existence of flow features with high predictive power for bulk TCP and VoIP traffic.
We show that the new agents offer a significant increase in throughput of legitimate TCP traffic for many choices of host density.
\end{abstract}

@@

However, the internet is always in flux.
The constant deployment of new services and protocols means that traffic is \emph{non-stationary} and evolving.
This ranges from subtle variation, such as new TCP congestion control algorithms \cite{rfc8312}, to the rapid rollout \cite{DBLP:conf/pam/RuthPDH18} of future protocols like QUIC \cite{DBLP:conf/sigcomm/LangleyRWVKZYKS17}.
%?? What happens when we roll out new protocols? Particularly those which focus on obfuscation? Behaviours become completely opaque, and ossification works against us.
Reliance on a fixed network is at odds with reality, when QUIC causes a large proportion of (traditionally) constant bitrate UDP traffic to become congestion-aware.
Behaviours and signatures become harder still to extract if the main purpose of a protocol is obfuscation (again, as in QUIC).
%?? Also attacks evolve.
DDoS attacks themselves also evolve in both short and long timescales (i.e., a sequence of attack strategies vs.\@ unseen attacks) \cite{DBLP:conf/spw/KangGS16}.
Thus, we must look at defence models which can identify characteristics and behaviours of DDoS traffic \emph{automatically}.

%Network anomaly detection and intrusion detection/prevention are continually evolving problems, compounded by the partial, non-\emph{independent and identically distributed} (IID) view of data at each point in the network.
%Attacks and anomalous behaviours evolve, becoming more sophisticated or employing new vectors to harm a network or system's confidentiality, integrity, and availability without being detected \cite{DBLP:journals/comsur/BhuyanBK14}.
%These attacks and anomalies have measurable consequences and symptoms which allow a skilled analyst to infer new signatures for detection by misuse-based classifiers, but unseen attacks may only be defended against after-the-fact.
%This issue is inherent to \emph{misuse-} or \emph{signature-based} intrusion detectors, and it has been long-hoped that \emph{anomaly-based} detectors would surpass this by making effective use of statistical measures \cite{DBLP:journals/comsur/BhuyanBK14}.

%?? Then talk about ML. And its costs, say that its ADS/IDS weaknesses extend to DDoS prevention.
While \emph{machine learning} (ML) approaches seem suitable for this problem, there is a distinct lack of real-world deployments in the broader field of intrusion detection/prevention.
\Textcite{DBLP:conf/sp/SommerP10} posit that this stems from greater risks than other ML tasks: the high cost of errors, low tolerance for false positives; a lack of recent, high-quality training data; and traffic diversity and evolution.
It is immediately clear that these challenges also apply to DDoS prevention.
Learning is made harder still by the challenges encountered when learning from unlabelled or partial data.
%DDoS prevention exhibits many of these issues.
%?? Then Bam! Introduce RL.

\emph{Reinforcement learning} (RL) offers another perspective.
%?? Unclear explanation of RL here?
RL agents follow a \emph{policy} to interact with or control a system, while at the same time using observed performance to improve this policy online.
An agent then adaptively reacts to threats by acting as a feedback loop for network optimisation---this allows us to overcome some of these difficulties by monitoring the consequences of each action in real-time.
We aim to prevent volume-based DDoS attacks with the aid of RL-based techniques, while bringing to light the flexibility of these techniques in the security domain.
%Whether it takes direct control of the network, or is used indirectly to optimise a key part of another system, more powerful `deep' RL techniques (and well-founded action spaces) aren't yet well explored for network IDS/IPS.
%These range from more modern training algorithms \cite{DBLP:journals/corr/SchulmanWDRK17, DBLP:conf/icml/SchulmanLAJM15}, to evolutionary strategies \cite{DBLP:journals/corr/SalimansHCS17, DBLP:journals/corr/abs-1802-08842}, hierarchical action composition \cite{DBLP:journals/corr/abs-1710-09767}, and competitive multi-agent learning \cite{DBLP:journals/corr/abs-1710-03748}.
%?? This raises the question: how do we do this effectively? Agents need state and reward info, which can be a lot of data transfer. THis gets worse (hint: we have the answer...)
This introduces a new set of problems: what effective state/action spaces should look like, how state and reward information should be efficiently propagated, and how this can be achieved on commodity hardware.

%?? Discuss the most important conclusions before the outline.
We find that the existing RL work for this task fails to account for congestion-aware traffic (i.e., TCP, QUIC) and environments with high host density at ingress points.
To remedy this, our agents make throttling decisions per-source, and show how an action model built on domain-specific knowledge can increase performance.
This requires additional engineering: updating RL agents from multiple traces per timestep, smart work scheduling, and an architecture featuring (but minimising) the use of software-defined networks (SDN).
%In tandem with the development and evaluation of an effective state space and model, we provide the design of a second model inspired by past work on algorithmic DDoS prevention, as an example of the integration of domain-specific knowledge.
This improves substantially upon the state-of-the-art when acting upon most internet traffic (i.e., congestion-aware protocols), and we show that the use of domain knowledge achieves excellent performance at high host density.
Crucially, the models are protocol- and content-agnostic to offer future-proofing against the rollout of future protocols.
%?? Also algorithmic enhancements such as multiple actions per timestep, 
%?? PROTOCOL-AGNOSTIC -- HOW WILL THESE THINGS COPE WITH QUIC ET AL.?!?!
Briefly, this paper contributes:
\begin{itemize}
%	\item A DDoS mitigation system based on direct-control reinforcement learning, designed for deployment in real-world software-defined networks (\cref{sec:environment-and-rl-algorithm,sec:acting-on-individual-sources}).
%	\item Important weaknesses and flaws in the past design and evaluation of similar techniques \cite{DBLP:journals/eaai/MalialisK15}. We offer a deconstruction and empirical study on their formulation, flaws, and risk factors concerning different traffic distributions---in addition, we demonstrate modifications enabling deployment in real-world software-defined networks (\cref{sec:performance-in-an-emulated-environment}).
	\item Two source-level granularity models for RL-driven DDoS prevention (\emph{Instant} and \emph{Guarded} action models) and suitable features (\cref{sec:ddos-mitigation-with-per-flow-reinforcement-learning}).
	\item A description of the system architecture which supports the real-world deployment of such agents (\cref{sec:system-architecture}).
%	\item An investigation of suitable features for this task (\cref{sec:rethinking-the-state-space}).
	\item Reactive simulations of HTTP and VoIP web-server traffic (\cref{sec:a-new-normal}).
	\item A performance comparison of different RL algorithms within these agents, and against the state-of-the-art in RL-based DDoS mitigation (\cref{sec:the-results-of-doing-so}).
%	\item An empirical evaluation of our new models against the state-of-the-art in RL-based DDoS mitigation (\cref{sec:the-results-of-doing-so}).
\end{itemize}

@@

\emph{Distributed denial of service} (DDoS) attacks are concentrated efforts by many hosts to reduce the availability of a service, to inflict financial harm or as an act of vandalism.
Attackers achieve this by exploiting system or application behaviour in \emph{semantic attacks} (e.g., \emph{SYN flooding attacks}), or by overwhelming their target through requests or traffic (\emph{volume-based attacks}) \cite{DBLP:conf/imc/JonkerKKRSD17}.
Hosts often participate unwillingly, typically having been recruited into a \emph{botnet} by malware infection to be orchestrated from elsewhere \cite{DBLP:conf/uss/AntonakakisABBB17}.

Volume-based attacks are the most relevant to our work, with two varieties being especially prevalent.
\emph{Amplification attacks} exploit services who send large replies in response to small requests, with UDP-based services like DNS and NTP seeing most abuse \cite{DBLP:conf/ndss/Rossow14, DBLP:conf/uss/KuhrerHRH14}.
Malicious hosts send many small requests, spoofed using the victim's address, causing many large replies to be sent to the intended target---increasing a botnet's throughput while keeping participants anonymous.
\emph{Transit-link/link-flooding attacks} have been the subject of recent attention, wherein malicious traffic is forwarded across core links needed to reach a target (but not to the target itself) \cite{DBLP:conf/sp/KangLG13, DBLP:conf/esorics/StuderP09}.

@@

Computing the approximate value of every action available in the current state forms the basis of a policy.
To encourage early exploration, actions with maximal value can be chosen each time except when taking random actions with probability $\epsilon$---an \emph{$\epsilon$-greedy policy}).
Each action $a \in \operatorname{A}(s)$ has probability $\pi(a|s,\wvec{})$ of being chosen:
\begin{equation}
	\pi(a|s,\wvec{}) =
	\begin{cases}
	1 - \epsilon + \epsilon/{|}{\operatorname{A}(s)}{|} & \text{if $a = \operatorname*{argmax}\limits_{a'} \acval{s}{a'}{\wvec{}}$} \\
	\epsilon/{|}{\operatorname{A}(s)}{|} & \text{otherwise}
	\end{cases}
\end{equation}
\Cref{fig:tilecode-select} shows how the value of each action is computed (and which action would be chosen by a greedy policy).

@@

\section{DDoS Mitigation with Per-flow Reinforcement Learning}\label{sec:ddos-mitigation-with-per-flow-reinforcement-learning}
We hypothesise that per-flow agent designs will offer better accuracy than current RL-based agents for DDoS mitigation.
We describe and justify our approach and algorithmic improvements, presenting two action models, one of which draws on domain knowledge introduced by SPIFFY \cite{DBLP:conf/ndss/KangGS16}.

@@

\subsection{Algorithm}
To make decisions and learn cheaply with low latency, we use \emph{semi-gradient Sarsa with tile coding} (\cref{eqn:sg-sarsa}).
Exploration is introduced by making use of $\epsilon$-greedy action selection, linearly annealing $\epsilon$ to 0 over time.
Each agent has its own internal parameter vector $\wvec{}$.

@@

\subsubsection{Action rate}
The algorithm is adapted to prioritise rapid response to changes in network state and to visit as many state-to-state transitions as possible to learn quickly.
Agents thus make many decisions per timestep.
We update $\wvec{}$ using each decision made and the reward signal from the relevant destination.
As exploration still occurs for each action, $\epsilon$ must be reduced multiple notches every timestep.
In turn, we increase the annealing window for $\epsilon$ by a factor of \num{2.67} so as to preserve exploration over time, by accounting for the greater volume of decisions being made.

\subsubsection{Per-tile updates}
To adjust the value of each activated tile with a different magnitude and direction, we compute a different temporal difference value \emph{for each tiling}.
This corrects tiles which contribute too much or too little value without disturbing estimates which are almost correct.
While we make use of all tiles' action value estimates when making decisions, each tiling is updated using only its own contribution, allowing us to increase $\alpha$ without divergence.
Compared to the standard algorithm, effective policies are reached sooner.

@@

		\end{tikzpicture}
	}
	\caption{
		Global state selection for a flow between an external host and server $s_0$ which passes over Agent 1.
		Nodes in the path taken are filled in blue, and link load measurements which are used for decisions are indicated with a thick blue line.
		\vspace{-1em}
		\label{fig:global-state-path}
	}
\end{figure}

@@

\Cref{fig:udp-feature-plots} shows the effectiveness of per-flow features for UDP (resp.\ \cref{fig:tcp-cap-feature-plots} for TCP), on a single-destination topology (\cref{sec:single-dest}) with $n=2$ hosts per egress point averaged over 10 episodes.
The plots demonstrate that different protocols and traffic classes are best defended by different features---every feature presented has value in a complete model.
Packet-level and per-window statistics are at their best when combined with the last action for both UDP-like and TCP-like traffic (omitted).
All features converge to their highest-observed performance within around \num{4000} timesteps.
In general, some of the most effective features are global state, mean IAT, mean inbound packet size and $\Delta$ rates.
The exclusion of features such as source/destination ports or protocol numbers is a deliberate choice.
If \emph{QUIC} (or a similar protocol) were to become ubiquitous, then these fields would have no correlation with the class of traffic a flow contains.

@@

\newcommand{\arrload}[2]{\operatorname{load}^{#2}_{t}(#1)}
\newcommand{\uload}[1]{\arrload{#1}{\uparrow}}
\newcommand{\dload}[1]{\arrload{#1}{\downarrow}}
\newcommand{\bload}[1]{\arrload{#1}{\updownarrow}}
\newcommand{\cond}[2]{\operatorname{c}_{#1,t}#2}
%\fakepara{Reward function directionality}
%The reward functions, as defined, do not take traffic direction into account.
%We redefine these to identify overload states using both upstream and downstream loads, while allowing customisation of which direction is chosen for protection.
We denote the upstream, downstream and combined loads at a node $s$ as $\uload{s}, \dload{s}, \bload{s}$ respectively.
Each destination node $s$ generates a reward signal, $R_{s,t}$, at every timestep $t$.
Assuming each destination has access to some classification function $g(\cdot)$ which estimates the volume of legitimate traffic received, and expects to receive $\mathit{traffic}_s$, we may compute the reward:
\begin{subequations}
%	\vspace{-0.5cm}
	\newcommand{\load}[1]{\operatorname{load}_{t}(#1)}
	\begin{gather}
	\cond{s} = [\max(\uload{s}, \dload{s}) > U_s],\\
	R_{s,t} = (1 - \cond{s}) \frac{g(\arrload{s}{-})}{\mathit{traffic}_s} - \cond{s},\label{eqn:reward-rtx}
	\end{gather}
	replacing $\arrload{s}{-}$ in \cref{eqn:reward} with whichever directional load is prioritised according to the traffic characteristics of the deployment environment, where $\cond{s}$ represents the ``overloaded'' condition at destination $s$.\label{eqn:reward}
\end{subequations}
We choose $\uload{\cdot}$ for our UDP-based models and $\dload{\cdot}$ for HTTP, and suggest $\bload{\cdot}$ for general deployment or heterogeneous traffic.
These choices reflect where the bulk of transmitted bytes in each traffic model are observed (or the lack of this knowledge).

While our use and definition of $g(\cdot)$ appears nebulous, there are many ways to infer this quantity in practice.
End-host servers may make use of canary flows or active measurement schemes, and can employ quality-of-experience metrics in the case of VoIP services such as lost packets, reorderings, and jitter.
ASes and transit networks may make use of reports received from downstream networks, i.e. over the \emph{DDoS Open Threat Signalling} (DOTS) protocol \cite{ietf-dots-use-cases-17}.
We leave the use of such data as future work.
Even without online rewards, a well-trained agent needs only to follow the greedy policy it has learned from offline (simulated) training.

@@

\subsubsection{Guarded control}
\Textcite{DBLP:conf/ndss/KangGS16} suggest that bot attack flows cannot scale up to match an increase in available bandwidth.
We apply their observations, constraining how an agent treats each flow via simple finite state automata: restricting $p \in \left\{ 0.00, 0.05, 0.25, 0.50, 1.0 \right\}$.
The action set is then simply to \emph{maintain}, \emph{increase}, or \emph{decrease} $p$ for a flow in single steps.

We choose these potential values for $p$ to add complete filtering to a steady progression of rate-limits (\SI{25}{\percent} increments for UDP traffic).
The outlier, $p=0.05$, corresponds to roughly a \SI{50}{\percent} rate reduction for TCP flows in our test topology.
This uneven spread of choices for $p$ allows light and heavy rate reduction to be applied to both congestion-aware and congestion-unaware traffic as required.
As each agent now requires the capability to plan ahead, we require a discount factor $\gamma \ne 0$, allowing the value of future states to influence state-action value updates.
We found the setting $\gamma = 0.8$ to be the most effective choice during exploratory testing.

@@

%?? Here is where I would put my ramblings about system architecture...
We present our design of a system which supports the effective real-world deployment of RL-based DDoS mitigation.
\Cref{fig:sys-arch} displays this, separating system elements which are local to each agent from those which reside elsewhere in the network.
We operate each agent as a \emph{virtual function} (VF) adjacent to a software-defined switch.
Software-defined networks allow a controller (or authorised hosts) to install actions, forwarding rules and logic into a switch at runtime.
Moreover, networks of this kind more naturally enable the future relocation and easy installation of learners.
Agent VFs communicate with these co-hosted switches to install probabilistic packet-drop rules.
We describe the main purpose and operation of each module within an agent's VF, and discuss techniques to make deployment more efficient using existing technologies.

@@

We use OpenFlow groups to simplify control messages: premade tables with permitted levels of packet drop.
This saves some overhead compared to using experimenter/extension headers.
Flows are automatically given a group with the default level of packet drop (according to the chosen agent design), meaning that switches don't need to refer to a controller or the agent VF.

@@

\section{Conclusions and Future Work}
Through this paper, we have introduced reinforcement learning and its relevance to network intrusion prevention.
We believe that the ability to learn network control as an online feedback loop makes it particularly effective in modern networks.

>>>>

\hypersetup{
	colorlinks,
	citecolor=black,
	filecolor=black,
	linkcolor=black,
	urlcolor=black,
	pdftitle={\mytitle{}},
%	pdfauthor={Kyle A. Simpson, Simon Rogers, Dimitrios P. Pezaros}
	pdfauthor={Anonymous Giraffe, Anonymous Ibex, Anonymous Badger}
}
\newcommand*{\email}[1]{\href{mailto:#1}{\nolinkurl{#1}} }

@@

\title{\mytitle{}}
\author{
%	Kyle~A.~Simpson\thanks{This work was supported by the Engineering and Physical Sciences Research Council [grant number EP/M508056/1].},
%	Kyle~A.~Simpson,
%	Simon~Rogers,
%	Dimitrios~P.~Pezaros,~\IEEEmembership{Senior~Member,~IEEE,},\\
%	School of Computing Science, University of Glasgow, Glasgow, G12 8RZ, UK\\
%	University of Glasgow, Glasgow, Scotland,\\
%	\email{k.simpson.1@research.gla.ac.uk}
Anonymous Giraffe,
Anonymous Ibex,
Anonymous Badger\\
Unnamed Department, Nowhere\\
\email{giraffe.a@unnamed.com}
}

@@

Reinforcement learning (RL) agents manage and monitor system health in an online manner.
Unsupervised, online learning is a key part of overcoming this detection problem.
Transfer of training data and execution can be expensive, hampering deployment.
In this paper, we show how to deploy such RL agents in a typical network at low cost through a novel architecture.
Our two agent designs make better use of flow information to advance the state-of-the-art in using RL for DDoS mitigation.
Both of these act per flow, in a protocol-agnostic manner for any network topology.
This is supported by an in-depth investigation of feature suitability and empirical evaluation.

Our results show that flow features effective on bulk TCP and VoIP traffic, combined with our agent designs, protect \SI{8}{$\!\times$} more TCP traffic than no defence at high host density.
This is a \SI{2.5}{$\!\times$} improvement over RL's state-of-the-art in DDoS defence.

%(\SI{2.5}{$\!\times$} improvement over Marl, \SI{8}{$\!\times$} more traffic than no defence system)
\end{abstract}

@@

However, the Internet is always in flux.
The constant deployment of new services and protocols means that traffic is \emph{non-stationary} and evolving.
This ranges from subtle variation, such as new TCP congestion control algorithms \cite{rfc8312}, to the rapid rollout \cite{DBLP:conf/pam/RuthPDH18} of future protocols like QUIC \cite{DBLP:conf/sigcomm/LangleyRWVKZYKS17}.
%?? What happens when we roll out new protocols? Particularly those which focus on obfuscation? Behaviours become completely opaque, and ossification works against us.
Reliance on a fixed network is at odds with reality, when QUIC causes a large proportion of (traditionally) constant bitrate UDP traffic to become congestion-aware.
Behaviours and signatures become harder still to extract if the main purpose of a protocol is obfuscation (again, as in QUIC).
%?? Also attacks evolve.
DDoS attacks themselves also evolve in both short and long timescales (i.e., a sequence of attack strategies vs.\@ unseen attacks) \cite{DBLP:conf/spw/KangGS16}.
Thus, we must look at defence models which can identify characteristics and behaviours of DDoS traffic \emph{automatically}.

While \emph{machine learning} (ML) approaches seem suitable for this problem, their deployment has been sparse in the broad field of intrusion prevention due to key risks: the high cost of errors; a low tolerance for false positives; and traffic diversity and evolution \cite{DBLP:conf/sp/SommerP10}.
These challenges also apply to DDoS prevention, made harder when learning from unlabelled or partial data.
\emph{Reinforcement learning} (RL) offers another perspective.
RL agents follow a \emph{policy} to control a system, using observed performance to improve this policy \emph{online}.
An agent then adapts to threats by acting as a feedback loop for network optimisation---allowing us to overcome some of these difficulties by monitoring the consequences of each action in real-time.
We aim to prevent volume-based DDoS attacks with the aid of RL techniques, while bringing to light their flexibility in the security domain.
This introduces a new set of problems: what effective state/action spaces should look like, how state and reward information should be efficiently propagated, and how this can be achieved on commodity hardware.

%?? Discuss the most important conclusions before the outline.
We find that the existing RL work for this task \cite{DBLP:journals/eaai/MalialisK15} fails to account for congestion-aware traffic (i.e., TCP, QUIC) and environments with high host density at ingress points.
To remedy this, our agents make throttling decisions per-source, and show how an action model built on domain-specific knowledge can increase performance.
This requires additional engineering: updating RL agents from multiple traces per timestep, smart work scheduling, and an architecture featuring the use of Software-defined Networks (SDN).
Our design minimises the use of (and thus attack surface introduced by) new technologies like SDNs, to increase compatibility with existing infrastructure and simplify operation.
These allow us to demonstrate how much value each feature offers when detecting attack flows.
%In tandem with the development and evaluation of an effective state space and model, we provide the design of a second model inspired by past work on algorithmic DDoS prevention, as an example of the integration of domain-specific knowledge.
This improves substantially upon the state-of-the-art when acting upon most internet traffic (i.e., congestion-aware protocols), and we show that the use of domain knowledge achieves excellent performance at high host density.
Crucially, the models are protocol- and content-agnostic to offer future-proofing against the rollout of future protocols.
%?? Also algorithmic enhancements such as multiple actions per timestep, 
%?? PROTOCOL-AGNOSTIC -- HOW WILL THESE THINGS COPE WITH QUIC ET AL.?!?!
Briefly, this paper contributes:
\begin{itemize}
%	\item A DDoS mitigation system based on direct-control reinforcement learning, designed for deployment in real-world software-defined networks (\cref{sec:environment-and-rl-algorithm,sec:acting-on-individual-sources}).
%	\item Important weaknesses and flaws in the past design and evaluation of similar techniques \cite{DBLP:journals/eaai/MalialisK15}. We offer a deconstruction and empirical study on their formulation, flaws, and risk factors concerning different traffic distributions---in addition, we demonstrate modifications enabling deployment in real-world software-defined networks (\cref{sec:performance-in-an-emulated-environment}).
	\item Two source-level granularity models for RL-driven DDoS prevention (\emph{Instant} and \emph{Guarded} action models) and suitable features (\cref{sec:ddos-mitigation-with-per-flow-reinforcement-learning}).
	\item A description of the system architecture which supports the real-world deployment of such agents (\cref{sec:system-architecture}).
%	\item An investigation of suitable features for this task (\cref{sec:rethinking-the-state-space}).
	\item Reactive simulations of HTTP and VoIP web-server traffic (\cref{sec:a-new-normal}).
	\item A performance comparison of different RL algorithms within these agents, and against the state-of-the-art in RL-based DDoS mitigation (\cref{sec:the-results-of-doing-so}).
%	\item An empirical evaluation of our new models against the state-of-the-art in RL-based DDoS mitigation (\cref{sec:the-results-of-doing-so}).
\end{itemize}

@@

\emph{Distributed denial of service} (DDoS) attacks may be broken down further---exploiting system or application behaviour in \emph{semantic attacks} (e.g., \emph{SYN flooding attacks}), or by overwhelming their target through requests or traffic (\emph{volume-based attacks}) \cite{DBLP:conf/imc/JonkerKKRSD17}.
Hosts often participate unwillingly, having been recruited into a \emph{botnet} by malware infection to be orchestrated from elsewhere \cite{DBLP:conf/uss/AntonakakisABBB17}.
Volume-based attacks are most relevant to our work, with two varieties being especially prevalent.
\emph{Amplification attacks} exploit services who send large responses to small requests, with UDP-based services like DNS and NTP seeing most abuse \cite{DBLP:conf/ndss/Rossow14, DBLP:conf/uss/KuhrerHRH14}.
Malicious hosts send many small requests, spoofed using the victim's address, causing the intended target to receive many large replies---increasing a botnet's throughput while keeping participants anonymous.
\emph{Link-flooding attacks} have been the subject of recent attention, wherein malicious traffic is forwarded across core links needed to reach a target (but not to the target itself) \cite{DBLP:conf/sp/KangLG13, DBLP:conf/esorics/StuderP09}.

@@

Computing the approximate value of every action available in the current state forms the basis of a policy.
To encourage early exploration, actions with maximal value can be chosen each time except when taking random actions with probability $\epsilon$---an \emph{$\epsilon$-greedy policy}).
Each action $a \in \operatorname{A}(s)$ has probability $\pi(a|s,\wvec{})$ of being chosen:
\begin{equation}
	\pi(a|s,\wvec{}) =
	\begin{cases}
	1 - \epsilon + \epsilon/{|}{\operatorname{A}(s)}{|} & \text{if $a = \operatorname*{argmax}\limits_{a'} \acval{s}{a'}{\wvec{}}$} \\
	\epsilon/{|}{\operatorname{A}(s)}{|} & \text{otherwise}
	\end{cases}
	\label{eqn:egreedy-pi}
\end{equation}
\Cref{fig:tilecode-select} shows how the value of each action is computed (and which action would be chosen by a greedy policy).

@@

\section{DDoS Mitigation with Per-flow Reinforcement Learning}\label{sec:ddos-mitigation-with-per-flow-reinforcement-learning}
We hypothesise that per-flow agent designs offer better accuracy than current RL-based agents for DDoS mitigation, because existing models apply one action over several flows.
Congestion-aware traffic is both more common and more vulnerable to this.
We describe and justify our approach and algorithmic improvements, presenting two different action models.

@@

\subsection{Algorithm}
To make decisions and learn cheaply with low latency, we use \emph{semi-gradient Sarsa with tile coding} (\cref{eqn:sg-sarsa}).
Exploration is introduced by making use of $\epsilon$-greedy action selection, linearly annealing $\epsilon$ to 0 over time (e.g., \num{10000} timesteps).
Each agent has its own internal parameter vector $\wvec{}$---giving its policy $\pi$.

@@

\subsubsection{Action rate}
The algorithm is adapted to prioritise rapid response to changes in network state and to visit as many state-to-state transitions as possible to learn quickly.
Agents thus make many decisions per timestep.
We update $\wvec{}$ using each decision made and the reward signal from the relevant destination.
As exploration still occurs for each action according to \cref{eqn:egreedy-pi}, and we take \numrange{2}{3} actions per timestep, $\epsilon$ must be reduced multiple notches every timestep.
We increase the annealing window for $\epsilon$ by a factor of \num{2.67} to preserve exploration over time, accounting for the greater volume of decisions being made.

\subsubsection{Per-tile updates}
To adjust the value of activated tiles with different magnitudes and directions, we compute a different temporal difference value $\delta_t$ \emph{for each tiling}.
Ordinarily, all tiles would change identically.
This corrects tiles which contribute too much or little value without disturbing estimates which are almost correct.
While we make use of all tiles' action value estimates when making decisions, each tiling is updated using only its own contribution, allowing us to increase $\alpha$ without divergence.
Compared to the standard algorithm, effective policies are reached sooner.

@@

		\end{tikzpicture}
	}
	\caption{
		Global state selection for a flow between an external host and server $s_0$ which passes over Agent 1.
		Nodes in the path taken are filled in blue, and link load measurements which are used for decisions are indicated with a thick blue line.
		\vspace{-1em}
		\label{fig:global-state-path}
	}
\end{figure}
\begin{figure}
	\centering
	\includegraphics[width=0.8\linewidth]{../plots/ftprep-cap-box}
	\vspace{-1cm}
	\caption{
		Learned performance of Instant Control agents when benign traffic is UDP-like, using only a single feature.
		%		Mean IAT, inbound packet sizes, and global state offer the best predictive performance.
		\label{fig:udp-feature-plots}
	}
\vspace{-1em}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=0.8\linewidth]{../plots/ftprep-tcp-cap-box}
	\vspace{-1cm}
	\caption{
		Learned performance of Instant Control agents when benign traffic is TCP-like, using only a single feature.
		%		All of the chosen features can offer a marked improvement over no protection at all.
		%		Global state and Mean IAT still offer the greatest improvement above baseline.
		\label{fig:tcp-cap-feature-plots}
	}
	\vspace{-1em}
\end{figure}

@@

\Cref{fig:udp-feature-plots} shows the effectiveness of per-flow features for UDP (resp.\ \cref{fig:tcp-cap-feature-plots} for TCP), on a single-destination topology (\cref{sec:single-dest}) with $n=2$ hosts per egress point averaged over 10 episodes.
``Baseline'' denotes an unprotected network.
The plots demonstrate that different protocols and traffic classes are best defended by different features---every feature presented has value.
Packet-level and per-window statistics are at their best when combined with the last action for both UDP-like and TCP-like traffic (omitted).
All features converge to their highest-observed performance within around \num{4000} timesteps.
In general, some of the most effective features are global state, mean IAT, mean inbound packet size and $\Delta$ rates.
The exclusion of features such as source/destination ports or protocol numbers is a deliberate choice.
If \emph{QUIC} (or a similar protocol) were to become ubiquitous, then these fields would have no correlation with the class of traffic a flow contains.

@@

\newcommand{\arrload}[2]{\operatorname{load}^{#2}_{t}(#1)}
\newcommand{\uload}[1]{\arrload{#1}{\uparrow}}
\newcommand{\dload}[1]{\arrload{#1}{\downarrow}}
\newcommand{\bload}[1]{\arrload{#1}{\updownarrow}}
\newcommand{\cond}[2]{\operatorname{c}_{#1,t}#2}
%\fakepara{Reward function directionality}
%The reward functions, as defined, do not take traffic direction into account.
%We redefine these to identify overload states using both upstream and downstream loads, while allowing customisation of which direction is chosen for protection.
We denote the upstream, downstream and combined loads at a node $s$ as $\uload{s}, \dload{s}, \bload{s} \in \mathbb{R}$ respectively.
Each destination node $s$ generates a reward signal, $R_{s,t}$, at every timestep $t$.
Assuming each destination has access to some classification function $g(\cdot)$ which estimates the volume of legitimate traffic received, and expects to receive $\mathit{traffic}_s$, we may compute the reward:
\begin{subequations}
%	\vspace{-0.5cm}
	\newcommand{\load}[1]{\operatorname{load}_{t}(#1)}
	\begin{gather}
	\cond{s} = [\max(\uload{s}, \dload{s}) > U_s],\\
	R_{s,t} = (1 - \cond{s}) \frac{g(\arrload{s}{-})}{\mathit{traffic}_s} - \cond{s},\label{eqn:reward-rtx}
	\end{gather}
	replacing $\arrload{s}{-}$ in \cref{eqn:reward} with whichever directional load is prioritised according to the traffic characteristics of the deployment environment, where $\cond{s}$ represents the ``overloaded'' condition at destination $s$.\label{eqn:reward}
\end{subequations}
We choose $\uload{\cdot}$ for our UDP-based models and $\dload{\cdot}$ for HTTP, and suggest $\bload{\cdot}$ for general deployment or heterogeneous traffic.
These choices reflect where the bulk of transmitted bytes in each traffic model are observed (or the lack of this knowledge).

While our use and definition of $g(\cdot)$ appears nebulous, there are many ways to infer this quantity in practice.
End-host servers may make use of canary flows or active measurement schemes, and can employ quality-of-experience metrics in the case of VoIP services such as lost packets, reorderings, and jitter.
ASes and transit networks may make use of reports received from downstream networks, i.e. over the \emph{DDoS Open Threat Signalling} (DOTS) protocol \cite{ietf-dots-use-cases-17}.
For testing, we use \emph{a priori} knowledge of which flows are legitimate; we leave the use of noisier real-world data as future work.
Even without online rewards, a well-trained agent needs only to follow the greedy policy it has learned from offline (simulated) training.

@@

\subsubsection{Guarded control}
\Textcite{DBLP:conf/ndss/KangGS16} suggest that bot attack flows cannot scale up to match an increase in available bandwidth.
We apply their observations, constraining how an agent treats each flow via simple finite state automata: restricting $p \in \left\{ 0.00, 0.05, 0.25, 0.50, 1.0 \right\}$.
The action set is then simply to \emph{maintain}, \emph{increase}, or \emph{decrease} $p$ for a flow in single steps.
This stops agents from wildly varying $p$ beyond their ability to correlate a host's response to its matching action.

We choose these potential values for $p$ to add complete filtering to a steady progression of rate-limits (\SI{25}{\percent} increments for UDP traffic).
The outlier, $p=0.05$, corresponds to roughly a \SI{50}{\percent} rate reduction for TCP flows in our test topology.
This uneven spread of choices for $p$ allows light and heavy rate reduction to be applied to both congestion-aware and congestion-unaware traffic as required.
As each agent now needs the capability to plan ahead, we require a discount factor $\gamma \ne 0$, a parameter which allows future value to influence value updates (\cref{eqn:sg-sarsa}).
We found $\gamma = 0.8$ to be the most effective choice during exploratory testing.

@@

%?? Here is where I would put my ramblings about system architecture...
We present our design of a system which supports the effective real-world deployment of RL-based DDoS mitigation.
\Cref{fig:sys-arch} displays this, separating system elements which are local to each agent from those which reside elsewhere in the network.
We operate each agent as a \emph{virtualised Network Function} (vNF) adjacent to a software-defined switch.
SDN allows a controller (or authorised hosts) to install actions, forwarding rules and logic into a switch at runtime.
Moreover, networks of this kind more naturally enable the future relocation and easy installation of learners.
Agent vNFs communicate with these co-hosted switches to install probabilistic packet-drop rules.
We describe the main purpose and operation of each module within an agent's vNF, and discuss techniques to make deployment more efficient using existing technologies.

@@

We use OpenFlow groups to simplify control messages: premade tables with permitted levels of packet drop.
This saves some overhead compared to using experimenter/extension headers.
Flows are automatically given a group with the default level of packet drop (according to the chosen agent design), meaning that switches don't need to refer to a controller or the agent vNF.

@@

\section{Conclusions}
In this paper we have put forth RL agent designs which act per-flow, describing the algorithmic modifications and engineering choices needed to make their deployment feasible.
Supporting this, we've demonstrated the value of the features these agents use as evidence.
Our evaluation against realistic normal traffic shows that our new agent designs considerably advance the state-of-the-art in RL-based DDoS prevention, protecting \SI{2.5}{$\!\times$} more TCP traffic than past RL work, with \emph{Guarded} agents showing the most promise for future evaluation.
